---
title: "Washington Cooperative Fish & Wildlife Research Unit"
author: "FY2020 Financial Report"
date: "August 25, 2020"
output: pdf_document
---

```{r include = FALSE} 

library(knitr)
library(scales)
library(ggplot2)
library(tidyr)
library(here)
library(tinytex)
library(reshape)
```

```{r setup, echo = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, results = 'asis', fig.width = 4, fig.height = 3.5)

projects <- read.csv(here("2020","ProjectData2020.csv"))
waiver <- read.csv(here("2020","WaiverUsage2020.csv"))
base <- read.csv(here("2020","BaseSpending2020.csv"))

Year <- "2020"
Year.m1 <- as.character(as.numeric(Year)-1)
Year.p1 <- as.character(as.numeric(Year)+1)
start.date <- paste("7/1/",Year.m1,sep="")
end.date <- paste("6/30/",Year,sep="")

coop.projects <- projects[c(projects$Coop == 1),]

totaldollars <- dollar(sum(coop.projects$TotalDollars))
newdollars <- dollar(sum(coop.projects$NewDollars))

totaldollars.WSU <- dollar(sum(coop.projects$TotalDollars[c(which(coop.projects$University == "WSU"))]))
newdollars.WSU <- dollar(sum(coop.projects$NewDollars[c(which(coop.projects$University == "WSU"))]))

totaldollars.UW <- dollar(sum(coop.projects$TotalDollars[c(which(coop.projects$University == "UW"))]))
newdollars.UW <- dollar(sum(coop.projects$NewDollars[c(which(coop.projects$University == "UW"))]))

funds.by.agency <- aggregate(coop.projects$TotalDollars,by=list(coop.projects$ParentAgency), FUN = sum)
funds.by.agency <- funds.by.agency[order(funds.by.agency$x,decreasing=TRUE),]
agency1 <- funds.by.agency[1,1]
agency2 <- funds.by.agency[2,1]
agency3 <- funds.by.agency[3,1]

total.funds.by.sponsor <- aggregate(coop.projects$TotalDollars,by = list(coop.projects$Sponsor,coop.projects$University),FUN=sum, drop = FALSE)

new.funds.by.sponsor <- aggregate(coop.projects$NewDollars,by = list(coop.projects$Sponsor,coop.projects$University),FUN=sum, drop = FALSE)

sponsor <- cbind(total.funds.by.sponsor,new.funds.by.sponsor[,3])
colnames(sponsor) <- c("Sponsor","University","Total Funds","New Funds")

Required <- c("WDFW-Fisheries","WDFW-Habitat","WDFW-Wildlife","WDNR","WDOE")
Required <- is.element(sponsor[,1],Required)
sponsor[intersect(which(Required == TRUE), which(is.na(sponsor$'Total Funds'==TRUE))),c(3,4)] <- 0
sponsor <- sponsor[complete.cases(sponsor), ]
sponsor$'Total Funds' <- dollar(sponsor$'Total Funds')
sponsor$'New Funds' <- dollar(sponsor$'New Funds')

```

## Total Project Funds

The WACFWRU Cooperative Agreement facilitated `r totaldollars.UW` in total project funds at the University of Washington and `r totaldollars.WSU` in total project funds at Washington State University in state fiscal year `r Year` (`r start.date` to `r end.date`), with `r newdollars.UW` in new project funds at University of Washington and `r newdollars.WSU` in new project funds at Washington State University. `r agency1` was the largest single funder, followed by `r agency2` and `r agency3`. 


```{r echo=FALSE}

kable(sponsor, caption = paste("Total funds and new funds by sponsor and university in FY",Year,sep=""))

waiver.full <- waiver[,c(1,2,ncol(waiver)-1,ncol(waiver))]

waiver <- melt(waiver.full, id=c("Agency","University"))
colnames(waiver) <- c("Agency","University","Year","Used")
total.waiver <- dollar(400000)

UW <- waiver[which(waiver$University == "UW"),-c(2)]
UW.curr.left <- dollar(400000 - sum(as.numeric(UW$Used[which(UW$Year == paste("FY",Year,sep=""))])))
UW.next.left <- dollar(400000 - sum(as.numeric(UW$Used[which(UW$Year == paste("FY",Year.p1,sep=""))])))

WSU <- waiver[which(waiver$University == "WSU"),-c(2)]
WSU.curr.left <- dollar(400000 - sum(as.numeric(WSU$Used[which(WSU$Year == paste("FY",Year,sep=""))])))
WSU.next.left <- dollar(400000 - sum(as.numeric(WSU$Used[which(WSU$Year == paste("FY",Year.p1,sep=""))])))

```

## Waiver Usage

At UW, `r UW.curr.left` of the `r total.waiver` waiver was unused by the end of FY`r Year`. As of the date of this report, `r UW.next.left` is still available for FY`r Year.p1`. At WSU, `r WSU.curr.left` was unused by the end of FY`r Year`, while `r WSU.next.left` is still available for FY`r Year.p1`.


```{r echo=FALSE, fig.height = 3, fig.width = 5}

ggplot(data = UW, aes(x=Agency, y = Used, fill=Year))+
geom_bar(position = "dodge",stat="identity")+
labs(y="UW $ Used", x = "Agency")+
ylim(0,200000)+
theme_bw()

```



Figure 1. Waiver used by agency and fiscal year at University of Washington (total available = $200,000 for WDFW, $100,000 for WDNR and WDOE).


```{r echo=FALSE, fig.height = 3, fig.width = 5}

ggplot(data = WSU, aes(x=Agency, y = Used, fill=Year))+
geom_bar(position = "dodge",stat="identity")+
labs(y="WSU $ Used", x = "Agency")+
ylim(0,200000)+
theme_bw()

```

Figure 2. Waiver used by agency and fiscal year at Washington State University (total available = $200,000 for WDFW, $100,000 for WDNR and WDOE).


```{r echo=FALSE, fig.height = 3, fig.width = 5}

balance <- dollar(base[nrow(base),ncol(base)])
spending <- dollar(base[nrow(base)-1,ncol(base)])
income <- dollar(71250)
base <- base[-c(nrow(base)),c(1,ncol(base))]

base[1:10,2] <- dollar(base[1:10,2])

```

## Base Spending

The WACFWRU base account balance was `r balance` at the end of state fiscal year 2019. Total spending on base funds was `r spending` in FY`r Year`, against total income of `r income`.

```{r echo=FALSE}
kable(base,caption = paste("Spending on base funds by category in FY",Year,sep=""))


```
