---
title: "Washington Cooperative Fish & Wildlife Research Unit"
author: "FY2021 Financial Report"
date: "September 16, 2021"
output: pdf_document
fontsize: 11pt
geometry: margin=1in
---

```{r include = FALSE} 

library(knitr)
library(scales)
library(ggplot2)
library(tidyr)
library(here)
library(tinytex)
library(reshape)
```

```{r setup, echo = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, results = 'asis', fig.width = 4, fig.height = 3.5)

projects <- read.csv(here("2021","ProjectData2021.csv"),header = TRUE)
waiver <- read.csv(here("2021","WaiverUsage2021.csv"),header = TRUE)
base <- read.csv(here("2021","BaseSpending2021.csv"),header = TRUE)

Year <- "2021"
Year.m1 <- as.character(as.numeric(Year)-1)
Year.p1 <- as.character(as.numeric(Year)+1)
start.date <- paste("7/1/",Year.m1,sep="")
end.date <- paste("6/30/",Year,sep="")

#keep only Coop projeects 
coop.projects <- projects[c(projects$Coop == "Yes"),]
coop.projects <- coop.projects[-c(which(is.na(coop.projects$StartDate)==TRUE)),]

#keep only projects that were active in the relevant year
start.after <- as.Date(coop.projects$StartDate, format = "%m/%d/%Y") > as.Date(end.date, format = "%m/%d/%Y") 
end.before <- as.Date(coop.projects$EndDate, format = "%m/%d/%Y") < as.Date(start.date, format = "%m/%d/%Y") 

annual.projects <- coop.projects[-c(union(which(start.after == TRUE),which(end.before == TRUE))),]

#calculate total new dollars per project (both rates) 
annual.projects$NewDollars_all <- annual.projects$NewDollars_1stRate+annual.projects$NewDollars_2ndRate
annual.projects$NewDollars_real <- annual.projects$NewDollars_all 

#don't allow negative new dollars 
annual.projects$NewDollars_real[c(which(annual.projects$NewDollars_real<0))]<- 0

#get total and new dollars by university   
totaldollars.WSU <- dollar(sum(annual.projects$Total_Dollars[c(which(annual.projects$PI_1_University == "WSU"))]))
newdollars.WSU <- dollar(sum(annual.projects$NewDollars_real[c(which(annual.projects$PI_1_University == "WSU"))]))

totaldollars.UW <- dollar(sum(annual.projects$Total_Dollars[c(which(annual.projects$PI_1_University == "UW"))]))
newdollars.UW <- dollar(sum(annual.projects$NewDollars_real[c(which(annual.projects$PI_1_University == "UW"))]))

#create sponsor (combine WDFW with division)
#annual.projects$cooperator <- annual.projects$Sponsor
#for(i in 1:nrow(annual.projects)){
#  if(annual.projects$Sponsor[i]=="WDFW"){
#  annual.projects$cooperator[i] <- paste(annual.projects$cooperator[i],annual.projects$Sponsor_Division[i],sep="-")}}

#get funds by sponsor   
funds.by.agency <-  aggregate(annual.projects$Total_Dollars,by=list(annual.projects$Sponsor), FUN = sum)
funds.by.agency <- funds.by.agency[order(funds.by.agency$x,decreasing=TRUE),]
agency1 <- funds.by.agency[1,1]
agency2 <- funds.by.agency[2,1]
agency3 <- funds.by.agency[3,1]

total.funds.by.sponsor <- aggregate(annual.projects$Total_Dollars,by = list(annual.projects$Sponsor,annual.projects$PI_1_University),FUN=sum, drop = FALSE)

new.funds.by.sponsor <- aggregate(annual.projects$NewDollars_real,by = list(annual.projects$Sponsor,annual.projects$PI_1_University),FUN=sum, drop = FALSE)

sponsor <- cbind(total.funds.by.sponsor,new.funds.by.sponsor[,3])
colnames(sponsor) <- c("Sponsor","University","Total Funds","New Funds")

Required <- c("WDFW","WDNR","WDOE")
Required <- is.element(sponsor[,1],Required)
sponsor[intersect(which(Required == TRUE), which(is.na(sponsor$'Total Funds'==TRUE))),c(3,4)] <- 0
sponsor <- sponsor[complete.cases(sponsor), ]
sponsor$'Total Funds' <- dollar(sponsor$'Total Funds')
sponsor$'New Funds' <- dollar(sponsor$'New Funds')

```

## Total Project Funds

The WACFWRU Cooperative Agreement facilitated `r totaldollars.UW` in total project funds at the University of Washington and `r totaldollars.WSU` in total project funds at Washington State University in state fiscal year `r Year` (`r start.date` to `r end.date`), with `r newdollars.UW` in new project funds at University of Washington and `r newdollars.WSU` in new project funds at Washington State University. `r agency1` was the largest single funder, followed by `r agency2` and `r agency3`. 

\vspace{16pt} 

```{r echo=FALSE}

kable(sponsor, caption = paste("Total funds and new funds by sponsor and university in FY", Year, ".", sep=""), align='lcrr')

waiver.full <- waiver[,c(1,2,ncol(waiver)-1,ncol(waiver))]

waiver <- melt(waiver.full, id=c("Agency","University"))
colnames(waiver) <- c("Agency","University","Year","Used")
total.waiver <- dollar(400000)

UW <- waiver[which(waiver$University == "UW"),-c(2)]
UW.curr <- dollar(sum(as.numeric(UW$Used[which(UW$Year == paste("FY",Year,sep=""))])))
UW.next <- dollar(sum(as.numeric(UW$Used[which(UW$Year == paste("FY",Year.p1,sep=""))])))

WSU <- waiver[which(waiver$University == "WSU"),-c(2)]
WSU.curr <- dollar(sum(as.numeric(WSU$Used[which(WSU$Year == paste("FY",Year,sep=""))])))
WSU.next <- dollar(sum(as.numeric(WSU$Used[which(WSU$Year == paste("FY",Year.p1,sep=""))])))

```

\vspace{16pt} 

\newpage

## Waiver Usage

At UW, `r UW.curr` of the `r total.waiver` waiver was used by the end of FY`r Year`. As of the date of this report, `r UW.next` has been allocated for FY`r Year.p1`. At WSU, `r WSU.curr` was used by the end of FY`r Year`, while `r WSU.next` has been allocated for FY`r Year.p1`.

\vspace{16pt} 


```{r echo=FALSE, fig.height = 3, fig.width = 5, fig.align = "center"}

ggplot(data = UW, aes(x=Agency, y = Used, fill=Year))+
geom_bar(position = "dodge",stat="identity")+
labs(y="UW $ Used", x = "Agency")+
ylim(0,200000)+
theme_bw()

```


\setlength{\leftskip}{0.5in}

\setlength{\rightskip}{0.5in}

Figure 1: Waiver used by agency and fiscal year at University of Washington (total available = $200,000 for WDFW, $100,000 for WDNR and WDOE).

\setlength{\leftskip}{0in}

\setlength{\rightskip}{0in}

\vspace{24pt} 

```{r echo=FALSE, fig.height = 3, fig.width = 5, fig.align = "center"}


ggplot(data = WSU, aes(x=Agency, y = Used, fill=Year))+
geom_bar(position = "dodge",stat="identity")+
labs(y="WSU $ Used", x = "Agency")+
ylim(0,200000)+
theme_bw()

```

\setlength{\leftskip}{0.5in}

\setlength{\rightskip}{0.5in}

Figure 2: Waiver used by agency and fiscal year at Washington State University (total available = $200,000 for WDFW, $100,000 for WDNR and WDOE).

\setlength{\leftskip}{0in}

\setlength{\rightskip}{0in}

\vspace{16pt} 

```{r echo=FALSE, fig.height = 3, fig.width = 5}

balance <- dollar(base[nrow(base),ncol(base)])
spending <- dollar(base[nrow(base)-1,ncol(base)])
income <- dollar(71250)
base <- base[-c(nrow(base)),c(1,ncol(base))]

base[1:11,2] <- dollar(base[1:11,2])

```

## Base Spending

The WACFWRU base account balance was `r balance` at the end of state fiscal year `r Year`. Total spending on base funds was `r spending` in FY`r Year`, against total income of `r income`.

\vspace{16pt} 

```{r echo=FALSE}
kable(base, caption = paste("Spending on base funds by category in FY", Year, ".", sep=""), align = "lr")


```
